### 2020-12-11 workflow/scripts/predictor_matrices.Rmd
predictor matrices rmd

---
title: "GLM Predictor Matrices"
subtitle: "Analyses Europe6 and Europe7"
author: "Cecilia Valenzuela"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  beamer_presentation:
    theme: "Pittsburgh"
    colortheme: "dove"
    fonttheme: "structurebold"
    highlight: pygments
classoption: "aspectratio=169"
fontsize: 8pt
tables: yes
---
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, echo = FALSE, results ='hide', message = FALSE}
library(knitr)
library(wpp2019)
library(tidyverse)
library(readr)
source("./scripts/flight_data.R")
```

## Description

Generation of the predictor matrices for the GLM model from different sources 
of information:

- Eurostats, European data and statistics, transport data: <https://ec.europa.eu/eurostat/web/transport/data/database>
- United Nations Population Division, World Population Prospects 2019: R library `wpp2019, data(pop)`
- Distance

```{r}
## Daily flight passengers

demes <- c(China = "CN", France = "FR", Germany = "DE", Italy = "IT", OtherEuropean = "OE", Spain = "ES")

# These datasets are dowload from eurostats 
intraUE_data <- "./data/avia_paincc.tsv"
extraUE_data <- "./data/avia_paexcc.tsv"

fligh_data <- get_flightData(intraUE_data, extraUE_data, countries = demes)

fligh_data_3i <- fligh_data %>%
  filter((MONTH %in% 1:3 & YEAR == 2020) | (MONTH == 12 & YEAR == 2019)) %>%
  mutate(EPOCH = case_when(
    MONTH == 2 ~ 2,
    MONTH == 3 ~ 3,
    TRUE ~ 1)) %>%
  group_by(GEO, PARTNER, EPOCH) %>%
  mutate(N = n(),
         EVALUE = sum(DVALUE/N)) %>%
  ungroup %>%
  # Transformations
  mutate(LVALUE = log(EVALUE + 1),
         SLVALUE = scale(LVALUE))

write(slv_flightMatrix_3i, "files/flight_matrix_3i.csv", sep = ",")
```

## Population counts

Population counts from World Population Prospects 2019, United Nations Population Division,  `wpp2019` R library. Values for **China**, **France**, **Germany**, **Italy**, **Other European** and **Spain**.

A predictor matrix for source population and a predictor matrix for destination population are generated, log transformed and standardized. Constant across demes.

\small
```{r, echo=FALSE}
data(pop)
population <- pop %>% filter(name %in% labels(demes)) %>% select(name, "2020") 
pop_eu_demes <- population %>% filter(name != "China") %>% summarise(s = sum(`2020`)) %>% pull(s)
pop_eu_total <- pop %>% filter(name == "Europe") %>% pull("2020")
population <- population %>%
  tibble::add_row(name = "OtherEuropean", `2020` = pop_eu_total - pop_eu_demes) %>%
  arrange(name)
```

```{r, echo=FALSE}
# Population source matrix
pop_source <- matrix(1, 6, 6) * population$`2020`
diag(pop_source) <- NA
vslpop_source <- scale(log(c(pop_source)))
slpop_source <- matrix(vslpop_source, 6, 6)
colnames(slpop_source) <- demes[population$name]
rownames(slpop_source) <- demes[population$name]

# Population destination matrix
pop_dest <- sweep(matrix(1, 6, 6), 2, population$`2020`, `*`)
diag(pop_dest) <- NA
vslpop_dest <- scale(log(c(pop_dest)))
slpop_dest <- matrix(vslpop_dest, 6, 6)
colnames(slpop_dest) <- demes[population$name]
rownames(slpop_dest) <- demes[population$name]

knitr::kable(list(round(slpop_source, 2), round(slpop_dest, 2)), booktabs = TRUE, caption = "Population counts source (left) and destination (right)")
```

\normalsize
Other European population count is computed as the difference between the Europe population and the sum of the populations from France, Germany, Italy and Spain.

## Distance between countries

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.