### 2021-02-21 reports/mt-thesis/methods.tex

\chapter{Material and Methods}

\section{The multytipe birth death model}

We have used the multitype birth-death phylodynamic model. Phylodynamics studies how phylogenetic trees are being generated and infers the population parameters behind that process. A multitype birth death process describes the evolution of an structured population. The process starts with one individual in one of the subpopulations (or demes) and it can give rise to another individual at rate $\beta$, die at death $\delta$, migrate to other deme or be sampled at rate $\phi$ to become part of the phylogeny.

Since the rate of the evolution of RNA virus, in specific SARS-CoV-2 is in the same scale than the epidemic spread, we can study the transmission dynamics by looking at the virus sequences.

The parameters that we will consider are the R0, the becoming uninfectious rate, the migration rates and the sampling proportion (proportion of no longer infectious individual that arer sampled and included in the phylogenetic tree).

Thus, the probability that we want to know is:

$P(T, P | D)$

This process describes the generation of the tree as part of the dynamics of the whole population, it is very rich and detailed in information. Therefore it is also possible to obtain information about the epidemic trajectory of the whole population. We can obtain the full sequence of events that describes the epidemic in each of the demes.

From the phylogenetic tree and the phylodynamics parameters we use Stochastic mapping to infer the ancestral traits and then to infer the epidemic trajectories.

$P(Tc | T, P)$

$P(E | Tc, P)$

\section{Infering epidemic trajectories}

Stochastic mapping of ancestral traits

Read paper and tim presentation

Stochastic mapping of epidemic trajectories

From the phylogenie with ancestral locations and the set of phylodynamic parameters we can simulate epidemic trajectories using an algorithm similar to Gillespie (with tau leaping and particle filtering? the subsampling). And most improtant, we can compute the probability of that specific trajectory given those parrameters and tree.

$P(E|Tc, P)$

Any trajectories without the events that are respresent in the tree has probability 0, so to avoid the simualtion of this trajectories, no time effcieint, we will enforce the events that arer represented in the tree (and will include a weigth). Also to avoid low probabilities trajectories, we will importance subsample one trajectory after a certain time of simulation according to the weigths detemrined by its probabilities.

\section{Modeling the early dynamics of SARS-CoV-2 in Europe}

In our case, we have 6 demes, one for each of the countries/geographic location. A birth represents a transmission event of covid-19, a death event is an individual becoming uninfectious by overcoming the disease, being isolated or dying. From the birth rate and death rate we can estuimate the R0, key vallue in the epidemic understanding and control. A migration event happens when an individual travels from one country to another, we know that is an important way of covid-19 spreading and we are interested on the dynamics.

We focus our analysis in the early spread in Europe of SARS-CoV-2, therefore we constrain the analysis to the period from the origin of the epidemic to March 8, when the first European lockdown in the Lombardy region started. During this time, for the European countries we expect an unimpeded spread of the virus that could be described by a constant R0 particular to each deme. However, the migration rates could have changed during these months due to the increase public awareness, so we will consider a different migration rate for origin to 23 jan, 23 jan to end of february and 1 to 8 of March. We assume a constant sampling proportion different for each of the demes, to account for difference in the sequencing efforts. The becoming uninfectious rate we will fixed in 36.5, assuming that one individual is infectious for a exponential time distirbuted with mean 10.

\section{Incorporating travel data}

To incorporatet travel data in the model we have use a GLM model described in Lemey et al 2009. We use a generalized linear model to describe the migration matrix KxK, where the rpedictors are the number of dayly flight from one country to another from EUROSTATS, the average distance between the centrroids of the country and the population sizes from origin and destinations. The predictors are log transformed, we add a pseudocount to make them all positive, and standardized as described in ..

$mij = c exp(\sum_{i = 1}^p \delta_i \beta_i xij)$

The coefficients beta describe the effect size of the predictors in the migrations rate and the delta coefficients act as a model selection variables, taking 0-1 values including orr not that specific predictor in the model.

\section{SARS-CoV 2 genome data set}

GISAID. Sequences from France, Germany, Italy, Spain and other European countries. Subsampling a sequence with probability equal to the probability of having a case that day in the country, inverseley weigthed by the probability of being sequenced that day.

\section{GLM predictos data}

EUROSTAT, countries considers. Distance and population datasets.

\section{Bayesian inference of phylodynamic parameters}

We use bayesian inference and Metropolis Hasting Monte Carlo algorith to infer our phylodynamics parameters and phylogeneit tree. We use the following priors

Clock rate and substitution model. Fixed to 8x10-4, HKY 4 gamma categories with priors...
BDMM-Prime epi parameterization:
R0
Become uninfectious rate
Sampling rates
Migration rates

\subsubsection{R0}
R0 constant except for China (and Italy? maybe not in the main analysis).

Prior Log Normal (0.8, 0.5) Median 2.2, 95IQR 0.8 to 5.9.
(Lai A. 2020 prior log normal (0,1) median 1.0 0.1-7.10)

- Main analysis: We fix China R0 to 2.7 (23 jan) 1.3 (10 feb) 0.8
- Use a decreasing paramater as Sarah with migration rate? But then we would need to add sequences.
- Or instead of fixing it we can put a strong informative prior.
- Allow Italy R0 to change March 1.

Meta-analyses R0:

R0 in China (3.14,95\%CI,2.40–4.09).\cite{Billah2020}

R0value 2.90 (2.32, 3.63
 2.9 (95\% CI: 2.1–4.5)\cite{Park2020}

Periods:
until 23 jan R0 2.9 
23 jan - 10 feb 1.1 dashboard cevo
10 feb - 8 mar 0.43 dashboard cevo

Become uninfectious rate same for all demes, fixed to 36.5.
Migrations among demes, 3 epochs.
Sampling proportion 0 before first sample. Upperbounded by number of cases.

Implementation BEAST 2.6.3 and BDMM-Prime package.
We run 5 chains of 10e7 iterations witth different seeds.

\section{Processing of epidemic trajectories}
In our case, for each set of parameters and tree we will simulate 10000 trajectories with an epsilon of 0.3 and select one of them with importance sampling.
We have one trtajectory for each step in the MCMC. We subsampled x.

\section{Data availability}
GISAID data table.

\section{Code availability and analyses reproducibility}
All the code is availables at ...
We have implement the full workflow with Snakemake so its fully reproducible.