### 2021-03-19 reports/mt-thesis/discussion.tex
wip mt

\chapter{Discussion}
\todo[inline]{TODO discussion}

\begin{itemize}
\item Case counts discussion and reporting rates, second wave in Europe. Other estimations in other studies.

% Inferred population trajectories that account for all cases in the country. It is not an independent estimate since we are using informative priors for the sampling rates (upperbound computed as sequences/reported cases) and the subsampling scheme also includes the number of cases in the country (but that is a weaker link I think.\\

\item First introduction other studies.

\item Migration patterns other studies.

\item Discuss assumptions of the model, caveats and possible improvements.

\item Number of particles
\end{itemize}